---
title: "TI_DataPrep_old"
author: "Petra Borovska"
date: "1 července 2020"
output: html_document
---


html_chapters()

# General Set Up


clean environment and console

```{r clean_console}
cat("\014")
```



```{r clean_envir}
rm(list = ls())
```



Installing all necessary packages for data preparation 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# ## install packages
# install.packages("knitr")
# install.packages("devtools")
# install.packages("here")
# install.packages("tidyverse")
# devtools::install_github("hadley/emo")
# install.packages("viridis")
# devtools::install_github("mikabr/ggpirate")
# install.packages("afex")
# install.packages("bootES")
# install.packages("readr")


## load packages
library(knitr)
library(devtools)
library(here)
library(tidyverse)
library(emo)
library(viridis)
library(ggpirate)
library(afex)
library(emmeans)
library(multcomp)
library(bootES)
library(broom)
library(readr)



```







# Data Preparation




Data preparation, loding data set, basic adjustments


get current directory

```{r base_dir}

here()


```


list the files in the desired direcotry

```{r list_files}

here("dataSets")
list.files(here("dataSets"))  ## just double check if it's correctly saved

```




load the data for each data set

```{r load_data}
tr_sleep <- 
        read.csv(
                here(
                        "dataSets", 
                        "pilotA_sleep_tr.csv"
                )
        )

tr_wake <- 
        read.csv(
                here(
                        "dataSets", 
                        "pilotA_wake_tr.csv"
                )
        )


test_sleep <- 
        read.csv(
                here(
                        "dataSets", 
                        "pilotA_sleep_ts.csv"
                )
        )

test_wake <- 
        read.csv(
                here(
                        "dataSets", 
                        "pilotA_wake_ts.csv"
                )
        )
```



look at the data 


```{r show_data}

## show the data
# tr_sleep
# tr_wake
# test_sleep
# test_wake

```


clean data, delete not useful rows from each data set
Drop na only for testing, where the length of tibble is the same, not do for training, it varies
and it will drop also useful rows

```{r delete_empty}


test_sleep = test_sleep %>% drop_na()
test_wake = test_wake %>% drop_na()

```

```{r make_copy}

ts_sleep_R = test_sleep
ts_wake_R = test_wake

```

test_wake with the most data, I guess use that
need to get average from each pair
then I need to switch participants columns with rows

my columns are always:  
        pairType                _condition_token
        letterPos1              _condition_token
        letterPos2              _condition_token
        key_resp_test.corr      _condition_token    <- that might be different name with training
        key_resp_test.rt        _condition_token    <- also might be different for training


to calculate frequencies - total scores:
        premise   50
        1deg      20
        2deg      10
        anchor    10
        



```{r col_info}
colName_w_ts =
        ts_wake_R %>%
        colnames(.)
colNum_w_ts =
        ts_wake_R %>%
        ncol()

```



```{r summaryPerSubject}
# ts_wake_6_select =
#         ts_wake_R %>%
#         select(
#                 c(1:5)
#         )
#         
# 
# ts_wake_6_multiple =
#         ts_wake_6_select %>%
#         group_by(
#                 pairType_W_6
#         ) %>%
#         summarize(
#                 tibble(
#                 total_W_6 = sum(key_resp_test.corr_W_6), 
#                 rt_mean_W_6 = mean(key_resp_test.rt_W_6)
#                 )
#         ) %>%
#         add_column(max_W_6 = c(10, 20, 50, 10))

```



# Functions


write a funtion which gives me summary dfs, containing total number per pair type, mean rt per pair type
and total number of pair types per group


parameters that needs to be inserted:
        type of data set
        which columns from the data set, 5 columns for each participant - maybe create a range variables 
                for each participant
                

does not have to be insert as parameter, but need to be set up                
        vary pair type - it's always first column of new data set, so replace the column name
                it does not have to vary
        proportion column, again replace by number - it does not have to vary
        mean column - again replace by number - it does not have to vary


think about that when training - na s must be drop during the function call - when selection
        going on - so between step

For the function is good to make copy



```{r summaryPerSubject_function}


summary_subj_ts_f <- function(dataSet, s, e){
        
        s = as.integer(s)
        e = as.integer(e)

        dataSet_select = dataSet[, c(s:e)]
        dataSet_select = as_tibble(dataSet_select)
        
        
        origColNames = colnames(dataSet_select)[c(1, 4, 5)]
        
        colnames(dataSet_select) <- c("pairType", "letterPos1", "letterPos2",
                                      "key_resp_test.corr", "key_resp_test.rt")
        
        dataSet_multiple =
                dataSet_select %>%
                group_by(
                        pairType
                ) %>%
                summarize(
                        tibble(
                        total_corr = sum(key_resp_test.corr), 
                        mean_rt = mean(key_resp_test.rt)
                        )
                )

        colnames(dataSet_multiple) <- origColNames
        
        
        dataSet_multiple_t = 
                dataSet_multiple %>%
                gather(key = corr_cond_subj, value = value, 2:ncol(dataSet_multiple)) %>% 
                spread_(key = names(dataSet_multiple)[1],value = 'value')
        
        
        dataSet_multiple_t_head = slice_head(dataSet_multiple_t)
        dataSet_multiple_t_tail = slice_tail(dataSet_multiple_t)
        dataSet_multiple_t_less =
                dataSet_multiple_t_tail %>%
                dplyr::select(-1) %>%
                rename(anchor_rt = anchor, oneDegree_rt = oneDegree, 
                       premise_rt = premise, twoDegree_rt = twoDegree)

        oneRowDf = cbind(dataSet_multiple_t_head, dataSet_multiple_t_less)
        
        return(oneRowDf)
}





```


# Further Data Processing


Appliying the function.
First getting number of columns for each data set, 
so I know how many participants we have - total / 5.


Also getting name of the columns, so I know which tokens used = subject number.

```{r info_subj_number}

noCol_ts_W = ncol(ts_wake_R)
noCol_ts_S = ncol(ts_sleep_R)

noCol_ts_W    ## 25 columns / 5 = number of participants
noCol_ts_S    ## 20 columns / 5 = number of participants

namesCol_ts_W = colnames(ts_wake_R)
namesCol_ts_S = colnames(ts_sleep_R)

namesCol_ts_W
# tokens_W = 6, 9, 11, 15, 17

namesCol_ts_S
# tokens_S = 7, 8, 18, 19


```


Creating a df out of the function. 
Gives me total of corr responses per pair type and average rt per pair type.
Then make a proportion out of it and combine those df, into one bigger df.
Consider to write another function.


```{r apply_function_summary_subj_f}

## wake testing
subj_6 = summary_subj_ts_f(ts_wake_R, 1, 5)
subj_9 = summary_subj_ts_f(ts_wake_R, 6, 10)
subj_11 = summary_subj_ts_f(ts_wake_R, 11, 15)
subj_15 = summary_subj_ts_f(ts_wake_R, 16, 20)
subj_17 = summary_subj_ts_f(ts_wake_R, 21, 25)


## sleep testing
subj_7 = summary_subj_ts_f(ts_sleep_R, 1, 5)
subj_8 = summary_subj_ts_f(ts_sleep_R, 6, 10)
subj_18 = summary_subj_ts_f(ts_sleep_R, 11, 15)
subj_19 = summary_subj_ts_f(ts_sleep_R, 16, 20)

```


Now create a tibble out of those single outputs. 
First transpose rows and columns. <- that's already in the function
Then add together. <- also in the function


In the *** df_ts *** always update subject. A bit un elegant, 
but it's not that bad, I won't have more than 100 participants. 


```{r concat_tibble}

df_ts = rbind(subj_6, subj_9, subj_11, subj_15, subj_17, subj_7, subj_8, subj_18, subj_19)

```


Split the column with info on subj and conditon
Getting wide format. 
In order to get longer, transform data below.
Don't forget to change the levels into factors.


```{r totals_pairType}
premTot = 50
oneDegTot = 20
twoDegTot = 10
anchorTot = 10
```




```{r split_col_ts}

## complete df
df_ts_sep <- 
        df_ts %>%
        separate(
                corr_cond_subj,
                c("forDel1", "forDel2", "forDel3", "response", "condition", "subjNr")
        ) %>%
        dplyr::select(
                -c(1:4)
        ) %>%
        mutate(
                anchor_prop = anchor/anchorTot,
                oneDegree_prop = oneDegree/oneDegTot,
                premise_prop = premise/premTot, 
                twoDegree_prop = twoDegree/twoDegTot
        ) %>%
        rowwise() %>%
        mutate(
                inference_prop = mean(c(oneDegree_prop, twoDegree_prop))
        )


```

```{r excel_csv_df_ts_sep}

write_csv(
        df_ts_sep,
        here("outputFiles_R", "full_basicPairType_prop.csv")
)

```



```{r getting_longFormat}
recode_long <-
        df_ts_sep %>%
        pivot_longer(
                cols = c(anchor, oneDegree, premise, twoDegree),
                names_to = "cond",
                values_to = "proportion"
        )

## make factor out of cond
recode_long$cond = as.factor(recode_long$cond)

## verify
class(recode_long$cond)

```





For more detailed analysis on premise pair type only.
This is very ugly function

```{r copy_data}
copy_test_sleep = test_sleep
copy_test_wake = test_wake
```


```{r prepare_df_detailpairs}



summary_subj_ts_pairType_f <- function(dataSet_pair, l, u){
        
        l = as.integer(l)
        u = as.integer(u)
        
        oneSub_select <- 
        dataSet_pair %>%
        dplyr::select(
               l:u 
        )
        
        origColNames = colnames(oneSub_select)[c(1:4)]

        colnames(oneSub_select) <- c("pairType", "letterPos1", "letterPos2",
                                      "key_resp_test.corr")


        
        dataSet_pair_filter <-
                oneSub_select %>%
                filter(
                        pairType == "premise" | pairType == "oneDegree" |
                                pairType == "twoDegree"
                )
        
        BC <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "B" & letterPos2 == "C" |
                                letterPos1 == "C" & letterPos2 == "B"
                ) %>%
                summarise(
                        tibble(
                                BC = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)
        CD <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "C" & letterPos2 == "D" |
                                letterPos1 == "D" & letterPos2 == "C"
                ) %>%
                summarise(
                        tibble(
                                CD = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)
        DE <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "D" & letterPos2 == "E" |
                                letterPos1 == "E" & letterPos2 == "D"
                ) %>%
                summarise(
                        tibble(
                                DE = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)
        BD <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "B" & letterPos2 == "D" |
                                letterPos1 == "D" & letterPos2 == "B"
                ) %>%
                summarise(
                        tibble(
                                BD = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)
        CE <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "C" & letterPos2 == "E" |
                                letterPos1 == "E" & letterPos2 == "C"
                ) %>%
                summarise(
                        tibble(
                                CE = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)
        BE <-
                dataSet_pair_filter %>%
                filter(
                        letterPos1 == "B" & letterPos2 == "E" |
                                letterPos1 == "E" & letterPos2 == "B"
                ) %>%
                summarise(
                        tibble(
                                BE = sum(key_resp_test.corr)
                        )
                ) %>%
                gather(key = pair_cond_subj, value = value)


        pairType_df_ts = rbind(CD, CD, DE,
                               BD, CE, BE)

        pairType_df_ts_rename <-
                pairType_df_ts %>%
                add_column(
                pairLetter_subj = c(origColNames[1], origColNames[1], origColNames[1],
                                    origColNames[1], origColNames[1], origColNames[1]), 
                .before = "value"
        ) %>%
                separate(
                pairLetter_subj,
                c("forDel1", "condition", "subjNr")) %>% 
                dplyr::select(
                        -2
                        )

        
        return(pairType_df_ts_rename)
                
        
}


```


Getting info on tokens for detailed tibble with each premise pair.

```{r info_column_pair}

colNames_ts_sleep_pair = colnames(copy_test_sleep)
colNames_ts_wake_pair = colnames(copy_test_wake)


colNum_ts_sleep_pair = ncol(copy_test_sleep)    ## 10
colNum_ts_wake_pair = ncol(copy_test_wake)      ## 20

```




```{r apply_f}

subj_8_pair = summary_subj_ts_pairType_f(copy_test_sleep, 1, 4)
subj_7_pair = summary_subj_ts_pairType_f(copy_test_sleep, 6, 9)


subj_6_pair = summary_subj_ts_pairType_f(copy_test_wake, 1, 4)
subj_9_pair = summary_subj_ts_pairType_f(copy_test_wake, 6, 9)
subj_11_pair = summary_subj_ts_pairType_f(copy_test_wake, 11, 14)
subje_15_pair = summary_subj_ts_pairType_f(copy_test_wake, 16, 19)



```

```{r final_df_pair}

df_ts_pair= rbind(subj_6_pair, subj_9_pair, subj_11_pair, subje_15_pair, subj_7_pair, subj_8_pair)


```




